An eight-year-long low-frequency earthquake catalog for Southern Cascadia

### 2021-07-27 paper/agujournaltemplate.tex
Update figures

Once we have obtained a catalog for the period 2007-2009 using the data from the permanent seismic networks, we compare the LFE detections between the two catalogs: the FAME catalog (obtained with data from the FAME experiment) and the networks catalog (obtained from data from the permanent networks). As we may have many false detections, we try to eliminate some of them by assuming that LFEs present in both catalogs are always true detections. Then we define two thresholds: the first threshold is chosen such that at least half the LFE detections above the threshold in the FAME catalog are also in the network catalog, the second threshold is chosen such that at least half the LFE detections above the threshold in the network catalog are also in the FAME catalog. If we filter the catalogs and keep only detections above the thresholds, we are now confident that at least half of the LFE detections are true detections. As the number of stations recording may vary over time (especially for the network catalog), we keep only the LFE detections such that the associated cross correlation multiplied by the number of channels recording at that time is higher than the threshold. Thus, if there are few stations recordings at some time, the cross correlation must be higher for the LFE to be considered as a true detection. We then compared the normalized number of LFEs obtained with the two catalogs. Examples are given in Figures 5 and 6 for families B (inverted triangle in Figure 1) and A (square in Figure 1). Although the number of LFEs during each event may not be the same, we clearly see for family B that the timing of the events are the same for both catalogs. Family A is much more active but most events seem to be present in both catalogs as well. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/080413_07_023.eps}
\caption{Normalized cumulative number of LFEs over the period July 2007-July 2009 for the FAME catalog (red) and the network catalog (blue) for family B (inverted triangle in Figure 1).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/080401_05_050.eps}
\caption{Normalized cumulative number of LFEs over the period July 2007-July 2009 for the FAME catalog (red) and the network catalog (blue) for family A (square in Figure 1).}
\label{pngfiguresample}
\end{figure}

@@

In northern Cascadia, it has been observed that there is an increase in activity downdip of the plate boundary compared to up dip, both for tremor ~\cite{WEC_2011} and LFEs ~\cite{SWE_2019}. We want to verify whether we can see an increase of LFE activity for the downdip LFE families compared to the updip LFE families. Most families in the northern and central part of the study area are well aligned along the strike direction, however families in the southern part of the subduction zone are more distant from each other. We first look at five families aligned at latitude about 40.7N (group D, diamonds in Figure 1) and show the number of LFEs per day in Figure 8. We then look at three families aligned at latitude about 40.4N (group C, triangles in Figure 1) and show the number of LFEs per day in Figure 9. For both sets of families, LFE events are more frequent for the easternmost family than for the westernmost family, as was also the case in northern Cascadia ~\cite{SWE_2019}. It is more difficult to compare event size and swarm duration as the number of stations and the quality of the templates are different from each family, and may have a strong influence on the number of LFEs detected. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/set1_daily_LFEs.eps}
\caption{Daily number of LFEs over the period 2004-2011 for the five LFE families in group D (diamonds in Figure 1).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/set2_daily_LFEs.eps}
\caption{Daily number of LFEs over the period 2004-2011 for the three LFE families in group C (triangles in Figure 1).}
\label{pngfiguresample}
\end{figure}

The occurrence and amplitude of tectonic tremor can be modulated by small stress changes (about 1 kPa) associated with ocean or solid earth tides ~\cite{HOU_2015}. Sensitivity of LFE families activity to tidal stress changes has also been observed in northern Cascadia ~\cite{ROY_2015}, and on the San Andreas fault ~\cite{THO_2012}. We used calculated stresses from elastic deformations due to both water loads and bodily tides in the solid Earth along the Cascadia subduction zone (H. Houston, personal communication). The stress computation is done following the method described in ~\citeA{HOU_2015}. The resulting full stress tensors are projected to normal and shear (in the plate convergence direction) stresses on the interface, as well as mean stress (one-third the tensor trace). We then divided the values of tidal stress change into 20 stress bins and computed for each LFE family the number of LFEs occurring at a time when the value of the stress was in a given bin. We also computed what would be the expected number of LFEs occurring if the tidal stress changes have no influence on LFE activity. For most LFE families, we did not notice any evidence of influence of tidal stress on LFE activity. However, the LFE activity on family A seems to be correlated with tidal stress changes. Figure 10 shows the number of LFEs occurring during each stress bin (blue bars) and the expected number of LFEs (black line) for shear, normal, and mean stress. We also normalized the number of LFEs by the time during which the stress takes values in each bin. We note that LFE activity is triggered by negative shear stress and positive normal stress. This sensitivity of LE activity to tidal stress changes may explain while this family is much more active than the other families on the subduction zone.

>>>>

Once we have obtained a catalog for the period 2007-2009 using the data from the permanent seismic networks, we compare the LFE detections between the two catalogs: the FAME catalog (obtained with data from the FAME experiment) and the networks catalog (obtained from data from the permanent networks). As we may have many false detections, we try to eliminate some of them by assuming that LFEs present in both catalogs are always true detections. Then we define two thresholds: the first threshold is chosen such that at least half the LFE detections above the threshold in the FAME catalog are also in the network catalog, the second threshold is chosen such that at least half the LFE detections above the threshold in the network catalog are also in the FAME catalog. If we filter the catalogs and keep only detections above the thresholds, we are now confident that at least half of the LFE detections are true detections. As the number of stations recording may vary over time (especially for the network catalog), we keep only the LFE detections such that the associated cross correlation multiplied by the number of channels recording at that time is higher than the threshold. Thus, if there are few stations recordings at some time, the cross correlation must be higher for the LFE to be considered as a true detection. We then compared the normalized number of LFEs obtained with the two catalogs. Examples are given in Figures 5 and 6 for families A (square in Figure 1) and B (inverted triangle in Figure 1). Although the number of LFEs during each event may not be the same, we clearly see for family B that the timing of the events are the same for both catalogs. Family A is much more active but most events seem to be present in both catalogs as well. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/080401_05_050.eps}
\caption{Normalized cumulative number of LFEs over the period July 2007-July 2009 for the FAME catalog (red) and the network catalog (blue) for family A (square in Figure 1). The cumulative number of LFEs has been divided by the total number of LFEs so that both curves start at 0 and end at 1, in order to get a better match between the LFE bursts in both catalogs.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/080413_07_023.eps}
\caption{Normalized cumulative number of LFEs over the period July 2007-July 2009 for the FAME catalog (red) and the network catalog (blue) for family B (inverted triangle in Figure 1). The cumulative number of LFEs has been divided by the total number of LFEs so that both curves start at 0 and end at 1, in order to get a better match between the LFE bursts in both catalogs.}
\label{pngfiguresample}
\end{figure}

@@

In northern Cascadia, it has been observed that there is an increase in activity downdip of the plate boundary compared to up dip, both for tremor ~\cite{WEC_2011} and LFEs ~\cite{SWE_2019}. We want to verify whether we can see an increase of LFE activity for the downdip LFE families compared to the updip LFE families. Most families in the northern and central part of the study area are well aligned along the strike direction, however families in the southern part of the subduction zone are more distant from each other. We first look at three families aligned at latitude about 40.4N (group C, triangles in Figure 1) and show the number of LFEs per day in Figure 8. We then look at five families aligned at latitude about 40.7N (group D, diamonds in Figure 1) and show the number of LFEs per day in Figure 9. Finally, we look at five families aligned at latitude about 41.05N (group E, pentagons in Figure 1) and show the number of LFEs per day in Figure 10. For all three sets of families, LFE events are more frequent for the easternmost family than for the westernmost family, as was also the case in northern Cascadia ~\cite{SWE_2019}. It is more difficult to compare event size and swarm duration as the number of stations and the quality of the templates are different from each family, and may have a strong influence on the number of LFEs detected. \\

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/set1_daily_LFEs.eps}
\caption{Daily number of LFEs over the period 2004-2011 for the three LFE families in group C (triangles in Figure 1).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/set2_daily_LFEs.eps}
\caption{Daily number of LFEs over the period 2004-2011 for the five LFE families in group D (diamonds in Figure 1).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/set3_daily_LFEs.eps}
\caption{Daily number of LFEs over the period 2004-2011 for the five LFE families in group E (pentagons in Figure 1).}
\label{pngfiguresample}
\end{figure}

The occurrence and amplitude of tectonic tremor can be modulated by small stress changes (about 1 kPa) associated with ocean or solid earth tides ~\cite{HOU_2015}. Sensitivity of LFE families activity to tidal stress changes has also been observed in northern Cascadia ~\cite{ROY_2015}, and on the San Andreas fault ~\cite{THO_2012}. We used calculated stresses from elastic deformations due to both water loads and bodily tides in the solid Earth along the Cascadia subduction zone (H. Houston, personal communication). The stress computation is done following the method described in ~\citeA{HOU_2015}. The resulting full stress tensors are projected to normal and shear (in the plate convergence direction) stresses on the interface, as well as mean stress (one-third the tensor trace). We then divided the values of tidal stress change into 20 stress bins and computed for each LFE family the number of LFEs occurring at a time when the value of the stress was in a given bin. We also computed what would be the expected number of LFEs occurring if the tidal stress changes have no influence on LFE activity. For most LFE families, we did not notice any evidence of influence of tidal stress on LFE activity. However, the LFE activity on family A seems to be correlated with tidal stress changes. Figure 11 shows the number of LFEs occurring during each stress bin (blue bars) and the expected number of LFEs (black line) for shear, normal, and mean stress. We also normalized the number of LFEs by the time during which the stress takes values in each bin. We note that LFE activity is triggered by negative shear stress and positive normal stress. This sensitivity of LE activity to tidal stress changes may explain while this family is much more active than the other families on the subduction zone.
